\documentclass{article}


\usepackage[margin=0.6in]{geometry}
\usepackage{amssymb, amsmath, amsfonts}
\usepackage{mathtools}
\usepackage{physics}
\usepackage{enumerate}
\usepackage{array}
\newcommand{\Rl}{\mathbb{R}}
\newcommand{\f}[3]{#1\ :\ #2 \rightarrow #3}

\title{PMI 214 Notes}
\author{Sam Fleischer}
\date{November 8, 2016}

\begin{document}
    \maketitle

    \begin{itemize}
        \item Pyrethroid resistance is growing
        \item Malaria drugs' complete failure
        \item Selfish genes
        \item Homing Endonuclease, X-shredder, \textbf{Population Suppression} through extreme set-ratio distortion.  All offspring are males, then the male children also carry this mutation.
        \item CRISPR/Cas9: \textbf{Population replacement} of parasite susceptible with resistant.  Make a mosquito incapable of transmitting Malaria..
        \item How do we genetically modify mosquitoes?
        \begin{itemize}
            \item Transgenesis puts genes into mosquitoes
            \item All progeny have the genes as well
        \end{itemize}
        \item Mosquito ingests the gamete
        \item Then to a zygote
        \item Then to an ookinete
        \item Forms an oocyst - asexual reproduction
        \item cyst ruptures outside the midgut - releases sporozoites
        \item they reach the salivary glands
        \item A team at UCSD made transgenic mosquitoes which can't transmit Malaria - affects the ookinetes and the sporozoites
        \item Mendelian Inheritance prevents us from simply releasing them in to a village.  At best we'll have a low equilibrium.
        \item We need a gene drive - promote the spread of the trans gene to spread to the wild population
        \item Gene Drive Inheritance will lead to trans gene fixation.  This means getting away from standard Mendelian Inheritance
        \item A gene drive system utilizes homologous driven repair.  The gene drive targets a specific sequence of DNA in the mosquito - it cuts the DNA, and ``repairs'' it with the trans gene.  Forces offspring of mixed parents to be homozygote for the trans gene.
        \item The donor DNA includes the Control region, Effector region, and the Gene Drive
        \item ``Nice thing about DNA is that DNA is DNA, so...''
    \end{itemize}

\end{document}
















